Will Kamala Harris will win the 2024 Presidential election

### 2024-11-04 paper/paper.qmd
final push

Harris is currently leading the polls in Michigan with 49% of the vote and Trump with 47%, this is a rather slim margin, but the first in the discussed states where Harris is leading in the polls. In @fig-supportAM we can see that there was a significant jump for Harris and a small dip in October followed by a gradual increase leading up to the elections. When we look at the voting history of the state in the last two elections, the have voted for the same candidates as Arizona and Georgia, where they voted for Biden in 2020 and for Trump in 2016, with the voting percentages at 50.6% to 47.8% in 2020 and 47.5% to 47.3 in 2016 with respect to the winner.

Although unlike the previously mentioned states 2016 was the first time the state voted for the republican candidate since 1992. Considering Michigan is home to the largest black majority city in the United states, Detroit, and a large majority of union workers who often vote democratic as it is seen as the more union-friendly party between the two it is rather like that Harris will win the state @towin270.

@@

Pennsylvania is an interesting for several reasons, the first is that before Tim Walz was announced as Kamala Harris's running mate, one of the speculated candidates was Josh Shapiro, the current governor of Pennsylvania, which presidential candidates often consider from where their potential running mates are from as they tend to help with swing states nevertheless, she did not choose Shapiro as her running mate and this is no longer a factor.

Second, Similarly to Nevada and Michigan Pennsylvania Voted for Trump in 2016 with a 48.2% to a 47.5% victory, and for Biden in 2020 with a slightly more significant but not overwhelming victory at 50% to 48.8%. 2016 was the first time Pennsylvania voted for a republican candidate since 1988 @towin270.

Lastly, if we consider @fig-supportWP we see that harris has been polling at a consistent 48% since mid september which is somewhat similar to Georgia, while Pennsylvania seems to be leaning more democratic than republican than Georgia.@fig-supportGNN, they have both similarly increased their interest in Harris since the polling began until mid September which is when they both becan to plateau at their current polling numbers.

@@

The last swing state we will be discussing is Wisconsinng which we can see in our prediction table, @tbl-supporttable, has the high support at 48.38% for Harris from all this swing states we looked at. when we look at @fig-supportWP, se a small spike in early september for Harris, followed by a gradual stabilization at around 48% support for Harris similar to what our model shows. Now if we also take into ac count @fig-supportstate, we see that Wisconsin has the largest difference in support for Harris in comparison to all previously mentioned swing states.

Lastly, if we look at Wisconsin's voting history, the only time Wisconsin voted Republican since 1988 was in 2016 when the state gave its electoral votes for Trump with 47.2% of the vote to Clinton's 46.5% then in 2020 the state went back to blue with Biden winning the state, all things considered out of all states we took into account Harris is likely to win @towin270.

@@

Without considering the swing states and if we assume all the other states will vote as expected Kamala Harris has a slight lead of 226 electoral votes to Donald Trump's 219. Now, if we take the 43 combined votes of Arizona, Georgia, and North Carolina, and add it to Trumps tally. While allocate Nevada, Michigan, Pennsylvania, and Wisconsin's votes to Harris which sum up to 50 electoral votes. This totals to a Democratic victory for Kamala Harris at 276 votes to Donald Trump's 262, we have come to the conclusion that Kamala Harris is likely to win the upcomming United States election by 14 electoral votes. Below we have included an image of what we believe the electoral map will look after the election November 5th, @towin270:

@@

There are several limitations and weaknesses that could have affected out research and statistical analysis. First, the variable `numeric_grade` which is the variable that weighs pollster results for accuracy could have been used to determining the how we look at each pollster to make sure the pollster we chose was the most accurate of the most available, and could have included it in our model to increase accuracy of our predictions.Second, as the data from polls change often especially leading up to the election it is difficult to make realtime adjustments as new data becomes available. Next, Harris only announced her candidacy at the end of July, we did not have a very large time frame to take into account whenn looking at her polling data, as often presidential candidates begin campaigning much earlier than just a few months before the election. Furthermore, As we only created a Model for Harris and not Trump we had less to compare to, while we believe our results would not have changed significantly it would have been helpful to consider a model based around Donald Trump. Additionally we could have looked at the states recent voting history for their members of congress, both the senate , and the house of representatives, as well as the current political party holding power in the state, i.e. the party that the governor of each state is affiliated with. Lastly, in recent history elections have been rather close whether they were close in electoral votes or in the popular vote, pollster are hesitant to state that one candidate will win over another even if their research and data shows otherwise as they are afraid to lose their credibility for future elections if they turn out to be wrong or if their analysis is different than the actual results.

@@

As the election is quickly approaching there are not too many things we would be bale to consider for this election, nevertheless there are a few. First, While we did consider the historical voting patterns of each state, we have not done a rigorous analysis with that data, which we would be able to consider when analyzing a the topic in the future. As, poll results are inherently uncertain and are not absolute indicators of who will win the election, we could use confidence intervals in the future to demonstrate this uncertainty. Next, we could could add more variables into our analysis to create a stronger and more accurate model to help us predict the election with a higher degree of assurance. Lastly, we could consider intangibles more heavily in the future, including recent events, both local and worldwide, considering the socioeconomic status of voters and the voting patterns of certain ethnic groups and how the population and demographic has changed since the last election, as well as general sentiment towards the parties and their presidential candidates.

>>>>

Harris is currently leading the polls in Michigan with 49% of the vote compared to Trump’s 47%. This is a slim margin, but it is the first of the states we've discussed where Harris is leading in the polls. In @fig-supportAM, we see a significant jump in support for Harris, followed by a small dip in October, and then a gradual increase leading up to the election. Looking at the state’s voting history in the last two elections, Michigan has aligned with Arizona and Georgia, voting for Biden in 2020 and for Trump in 2016, with vote percentages at 50.6% to 47.8% in 2020 and 47.5% to 47.3% in 2016, respectively.
Unlike the previously mentioned states, however, 2016 was the first time since 1992 that Michigan voted for a Republican candidate. Considering Michigan is home to Detroit, the largest Black-majority city in the United States, and has a large population of union workers who often support the Democratic party due to its pro-union stance, it is likely that Harris will win the state (@towin270).

@@

Pennsylvania is an interesting case for several reasons. First, before Tim Walz was announced as Kamala Harris's running mate, one of the speculated candidates was Josh Shapiro, the current governor of Pennsylvania. Presidential candidates often consider the home states of potential running mates, as they can help with swing states. However, Harris ultimately did not choose Shapiro, so this is no longer a factor.
Second, similar to Nevada and Michigan, Pennsylvania voted for Trump in 2016 with a narrow margin of 48.2% to 47.5%, and for Biden in 2020 with a slightly larger but still close victory of 50% to 48.8%. Notably, 2016 was the first time Pennsylvania voted for a Republican candidate since 1988 (@towin270).
Lastly, if we look at @fig-supportWP, we can see that Harris has been polling consistently at around 48% since mid-September, which is somewhat similar to the trend in Georgia. However, Pennsylvania appears to be leaning more Democratic than Republican compared to Georgia. In @fig-supportGNN, both states show a similar increase in support for Harris since polling began, reaching a plateau in mid-September at their current polling numbers.

@@

The last swing state we will be discussing is Wisconsin, which, as we can see in our prediction table (@tbl-supporttable), has the highest support for Harris at 48.38% among all the swing states we examined. Looking at @fig-supportWP, we see a small spike in early September for Harris, followed by a gradual stabilization at around 48% support, consistent with our model’s projections. Additionally, in @fig-supportstate, Wisconsin shows the largest difference in support for Harris compared to all the previously discussed swing states.

Finally, if we consider Wisconsin's voting history, the only time it voted Republican since 1988 was in 2016, when the state gave its electoral votes to Trump with 47.2% of the vote, compared to Clinton's 46.5%. In 2020, the state returned to blue, with Biden winning. Considering all factors, Harris is likely to win Wisconsin and reach the 270 electoral votes needed to win the election (@towin270).

@@

Without considering the swing states and assuming all other states will vote as expected, Kamala Harris has a slight lead of 226 electoral votes to Donald Trump's 219. Now, if we allocate the 43 combined votes of Arizona, Georgia, and North Carolina to Trump and assign the votes of Nevada, Michigan, Pennsylvania, and Wisconsin to Harris, which total 50 electoral votes, the result is a Democratic victory for Kamala Harris with 276 votes to Donald Trump's 262. We conclude that Kamala Harris is likely to win the upcoming United States election by 14 electoral votes. Below, we have included an image of our projected electoral map for the November 5th election (@towin270):

@@

There are several limitations and weaknesses that could have affected our research and statistical analysis. First, the variable `numeric_grade`, which weighs pollster results for accuracy, could have been better utilized to determine how we evaluate each pollster. By ensuring that we selected the most accurate pollsters available and incorporating this into our model, we might have increased the accuracy of our predictions.

Second, since polling data often changes, especially leading up to an election, it is challenging to make real-time adjustments as new data becomes available. Additionally, because Harris only announced her candidacy at the end of July, we did not have a substantial time frame to analyze her polling data. Typically, presidential candidates start campaigning much earlier, which gives a more comprehensive data set for analysis.

Moreover, since we only created a model for Harris and not for Trump, we had limited comparisons. While we believe our results would not have significantly changed, having a model based on Trump could have provided a more balanced analysis. We also could have benefited from examining recent voting histories in each state for members of Congress, both in the Senate and the House of Representatives, as well as the current political party in power, such as the party affiliation of each state’s governor.

@@

As the election quickly approaches, there are not many additional factors we can consider at this stage. However, there are still a few areas for future improvement. First, while we did take into account the historical voting patterns of each state, we did not perform a rigorous analysis of that data. This is something we could explore more deeply in future studies. Additionally, because poll results are inherently uncertain and are not absolute predictors of the election outcome, we could use confidence intervals in the future to better illustrate this uncertainty.

Next, we could incorporate more variables into our analysis to create a stronger and more accurate model, which would help us make more reliable predictions. Lastly, we should consider intangibles more thoroughly, such as recent events both locally and globally, the socioeconomic status of voters, and the voting behavior of different ethnic groups. It would also be beneficial to analyze how population and demographic shifts since the last election might impact voting trends, as well as general public sentiment toward the political parties and their presidential candidates.
